NURM: a tool to locate numerical differences in pipelines

### 2019-06-04 paper.tex
edit some typos

Figure~\ref{fig:overview} shows an overview of our method. For
each tested condition, the pipeline is containerized with Docker, and its
interface is described with Boutiques~\cite{glatard2017boutiques}.
\reprozip~\cite{Chirigati2016} is added to one of the container images for
provenance tracking. 
Our tool starts by capturing temporary and multi write files in Condition 1 
(Fig.~\ref{fig:overview-capturing}). 
To do that, the tool first executes pipeline on Condition 1 to obtain process graph 
(Fig.~\ref{fig:overview-capturing} - (P)) and 
result files. Subsequently, it generates the list of processes that create these intermediary files. 
Using this list, the tool modifies docker image in Condition 1 and exectutes the pipeline 
again through it, which turns to produce results including temporary and overwritten
files (Fig.~\ref{fig:overview-capturing} - (C1)).
In the second step, the processes are labeled from 
the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
the tool builds docker image for Condition 2 by modifying pipeline processes that 
recognised from 
the process graph. Then, the modified docker image is executed in Condition 2 and 
each process is labeled right 
after the running by comparing checksum of the results obtained in Condition 1 
and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
We then label processes in 3 categories (Fig.~\ref{fig:overview-labelling} -
(L)) depending whether they create differences (represented in red) or not (green). 
Process labelling is done through hooks which are defined during the docker image modifying procedure, 
and incremental updates in Condition 2, 
until the end of the pipeline execution.

@@

\begin{figure}
\centering
  \includegraphics[scale=0.25]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{bash} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

For example, in the sample pipeline in 
Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} write a temporary file and 
also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
then overwritten by process \texttt{3\#fslmaths}. The list of these process are captured in 
Listing~\ref{listing:json_process}.
Using this list, we modify relevant processes in Condition 1 and then re-execute pipeline 
to store them in a backup folder.
Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
captured as the outputs in Condition 1.

@@

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In such 
situations, it is not possible to determine from the pipeline results 
whether the process created differences, removed some differences, or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing We identify dall the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with 
the arguments of a process that created differences. 
Differences are identified by comparing checksum of the output files for 
the same process in both conditions.
We developed an incremental approach that consists of the following steps:

@@

The four steps in Figure~\ref{fig:iterations} correspond to the 
steps of the labelling algorithm. 
At step one, process \texttt{1\#bet} is executed and then labeled 
as transparent (green) as it produced files without differences.
At step 2, \texttt{2\#bet2} 
is labeled as difference creator (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
Condition 1.
At step 3, after running process \texttt{3\#fslstats}, it labeled as 
transparent as it write files without differences.
At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
as the last process of the pipeline.
As a result of these 4 steps, the final process labelling is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red) 
and the other processes are transparent.
It should be noted that process that remove differrences are not recognised 
by this method.

@@

  \State /* Input: - a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
  \State /*\quad \quad \quad \quad - a distance threshold value; \texttt{threshold}
  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
  \State \texttt{\# Create a cluster for each tree in input list;}
  \State \texttt{C = [ [t${_1}$], [t${_2}$], \ldots , [t${_n}$] ]}
  \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
  \State \texttt{D} = \texttt{array(n, n)}
  \While{\texttt{len(C)} > \texttt{1}}  
  \State \texttt{\# Select the two nearest clusters}
  \For{\texttt{i=1} to \texttt{n}}
  \For{\texttt{j=i+1} to \texttt{len(C)}}
  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])} \tristan{how do you define the distance between two clusters?}
    \EndFor
  \EndFor
  \State \texttt{i, j} = $\argmin \left\{ \texttt{D[i][j]}, \ \texttt{i, j} \in \llbracket 1, \texttt{len(C)}\rrbracket^2 \right\}$
  \If{\texttt{D[i][j]} > \texttt{threshold}}
  \State \Return \texttt{C}
  \Else
  \State \texttt{\# Merge C[j] into C[i]}
  \State \texttt{C[i]} = \texttt{merge}(\texttt{C[i] and C[j]})
  \State \texttt{\# Remove C[j] from C}
  \State \texttt{remove(C[j], C)}
  \EndIf
  \EndWhile
%  \State Function edit\_distance (C${_1}$, C${_2}$)
%  \State EndFunction
\end{algorithmic}
\end{algorithm}

>>>>

Figure~\ref{fig:overview} shows an overview of our method. For
each tested condition, the pipeline is containerized with Docker, and its
interface is described with Boutiques~\cite{glatard2017boutiques}.
\reprozip~\cite{Chirigati2016} is added to one of the container images for
provenance tracking. 
Our tool starts by capturing temporary and multi write files in Condition 1 
(Fig.~\ref{fig:overview-capturing}). 
To do that, the tool first executes pipeline on Condition 1 to obtain process graph 
(Fig.~\ref{fig:overview-capturing} - (P)) and 
result files. Subsequently, it generates the list of processes that create these intermediary files. 
Using this list, the tool modifies docker image in Condition 1 and exectutes the pipeline 
again through it, which turns to produce results including temporary and overwritten
files (Fig.~\ref{fig:overview-capturing} - (C1)).
In the second step, the processes are labeled from 
the differences found in their outputs in the
two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
the tool builds docker image for Condition 2 by modifying pipeline processes that 
recognised from 
the process graph. Then, the modified docker image is executed in Condition 2 and 
each process is labeled right 
after the running by comparing checksum of the results obtained in Condition 1 
and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
We then label processes in 2 categories (Fig.~\ref{fig:overview-labelling} -
(L)) depending whether they create differences (represented in red) or not (green). 
Process labelling is done through hooks which are defined during the docker image modifying procedure, 
and incremental updates in Condition 2, 
until the end of the pipeline execution.

@@

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{bash} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

For example, in the sample pipeline in 
Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} write a temporary file and 
also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
then overwritten by process \texttt{3\#fslmaths}. The list of these processes are captured in 
Listing~\ref{listing:json_process}.
Using this list, we modify relevant processes in Condition 1 and then re-execute pipeline 
to store them in a backup folder.
Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
captured as the outputs in Condition 1.

@@

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In such 
situations, it is not possible to determine from the pipeline results 
whether the process created differences, removed some differences, or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with 
the arguments of a process that created differences. 
Differences are identified by comparing checksum of the output files for 
the same process in both conditions.
We developed an incremental approach that consists of the following steps:

@@

The four steps in Figure~\ref{fig:iterations} correspond to the 
steps of the labelling algorithm. 
At step one, process \texttt{1\#bet} is executed and then labeled 
as transparent (green) as it produced files without differences.
At step 2, \texttt{2\#bet2} 
is labeled as difference creator (red) as it produced files with differences 
from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
Condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
Condition 1.
At step 3, after running process \texttt{3\#fslstats}, it labeled as 
transparent as it write files without differences.
At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
as the last process of the pipeline.
As a result of these 4 steps, the final process labelling is: 
\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red) 
and the other processes are transparent.
It should be noted that processes that remove differrences are not recognised 
by this method.

@@

  \State /* Input: - a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
  \State /*\quad \quad \quad \quad - a distance threshold value; \texttt{threshold}
  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
  \State \texttt{\# Create a cluster for each tree in input list;}
  \State \texttt{C = [ [t${_1}$], [t${_2}$], \ldots , [t${_n}$] ]}
  \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
  \State \texttt{D} = \texttt{array(n, n)}
  \While{\texttt{len(C)} > \texttt{1}}  
  \State \texttt{\# Select the two nearest clusters}
  \For{\texttt{i=1} to \texttt{n}}
  \For{\texttt{j=i+1} to \texttt{len(C)}}
  \State \texttt{D[i][j]} = $\min \left\{ \texttt{edit\_dist(a, b)}: \ \texttt{a} \in \texttt{C[i]}, \texttt{b} \in \texttt{C[j]} \right\}$
    \EndFor
  \EndFor
  \State \texttt{i, j} = $\argmin \left\{ \texttt{D[i][j]}, \ \texttt{i, j} \in \llbracket 1, \texttt{len(C)}\rrbracket^2 \right\}$
  \If{\texttt{D[i][j]} > \texttt{threshold}}
  \State \Return \texttt{C}
  \Else
  \State \texttt{\# Merge C[j] into C[i]}
  \State \texttt{C[i]} = \texttt{merge}(\texttt{C[i] and C[j]})
  \State \texttt{\# Remove C[j] from C}
  \State \texttt{remove(C[j], C)}
  \EndIf
  \EndWhile
%  \State Function edit\_distance (C${_1}$, C${_2}$)
%  \State EndFunction
\end{algorithmic}
\end{algorithm}
